---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo = FALSE}
targets::tar_load(syst_all)
```


# systemizace

<!-- badges: start -->
<!-- badges: end -->

This repository contains R code to download and process official data on the organisation of public servants in the Czech central public administration - "systemizace služebních míst".

The code downloads, loads and processes the published excel files such that

- the data is in tidy format
- civil servants and employees are in one data frame
- pay grades (platové třídy) and management levels (představení) are correctly identified

See https://www.mvcr.cz/sluzba/clanek/systemizace-sluzebnich-a-pracovnich-mist.aspx

Data currently available from `r min(syst_all$year)` to `r max(syst_all$year)`.

For more info on data on public servant see [recent report](https://idea.cerge-ei.cz/zpravy/statni-zamestnanci-a-urednici-kde-pracuji-a-za-kolik) and [related repo](https://github.com/dan-bart/urednici_2021), plus an [older overview of available sources](https://petrbouchal.xyz/urednici/).

Core organised as a {targets} pipeline, with packages tracked by {renv}, so can be reproduced like so:

```r
renv::restore()
targets::tar_make()
```

Rendered `r Sys.time()`
